Variational Monte Carlo on Bosonic Systems

### 2018-04-05 project_1/article/article.tex
Spell check systems section

\section{Systems}
    To model the trapped bosonic gas particles we use the potential
    \begin{align}
        v(\vf{r})
        &=
        \begin{cases}
            \half m\omega^2r^2 & (\text{S}), \\
            \half m \bigl[
                \omega^2(x^2 + y^2) + \omega_z^2z^2
            \bigr] & (\text{E}),
        \end{cases}
    \end{align}
    where we can choose between a spherical (S) or an elliptical (E) harmonic
    oscillator trap.  In the latter potential the oscillatory frequency is
    different in the $z$-direction, i.e., $\omega_z \neq \omega$.  The full
    Hamiltonian of the system is given by
    \begin{align}
        H = \sum_{i = 1}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where the single particle one-body operator, $h(\vf{r}_i)$, is given by
    \begin{align}
        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
        + v(\vf{r}_i),
    \end{align}
    for equal masses, and the two-body interaction operator, $w(\vf{r}_i,
    \vf{r}_j)$, is
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        = \begin{cases}
            \infty & |\vf{r}_i - \vf{r}_j| \leq a, \\
            0 & |\vf{r}_i - \vf{r}_j| > a,
        \end{cases}
        \label{eq:two-body_interaction}
    \end{align}
    where $a$ is the hard sphere of the particle.  The trial wavefunction,
    $\Psi_T(\vf{r})$, we will be looking at is given by
    \begin{align}
        \Psi_T(\vf{r})
        &= \Phi_T(\vf{r})
        \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k) \\
        &= \Biggl(
            \prod_{i = 1}^N g(\alpha, \beta, \vf{r}_i)
        \Biggr)
        \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
        \label{eq:initial_trial_wavefunction}
    \end{align}
    where $\alpha$ is a variational parameter and
    \begin{align}
        \vf{r} = (\vf{r}_1, \vf{r}_2, \dots, \vf{r}_N, \alpha, \beta).
    \end{align}
    Here $g$ are the single particle wavefunctions given by
    \begin{align}
        g(\alpha, \beta, \vf{r}_i)
        = \exp\bigl[
            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
        \bigr] \equiv \phi(\vf{r}_i),
    \end{align}
    and $\Phi_T(\vf{r})$ is the \emph{Slater permanent} consisting of the $N$
    first single particle wavefunctions, and the correlation wavefunction, $f$,
    given by
    \begin{align}
        f(a, \vf{r}_j, \vf{r}_k)
        &=
        \begin{cases}
            0 & |\vf{r}_j - \vf{r}_k| \leq a, \\
            \Bigl(
                1 - \frac{a}{|\vf{r}_j - \vf{r}_k|}
            \Bigr) & |\vf{r}_j - \vf{r}_k| > a.
        \end{cases}
        \label{eq:correlation_wavefunction}
    \end{align}
    We will for brevity use the notation $\phi(\vf{r}_i) = \phi_i$ and $r_{jk} =
    |\vf{r}_j - \vf{r}_k|$. In practice \autoref{eq:two-body_interaction} is
    unnecessary as the situtation where $r_{jk} \leq a$ automatically yields a
    probability density of zero from \autoref{eq:correlation_wavefunction} thus
    rejecting the state completely.

>>>>

\section{Systems}
    To model the trapped bosonic gas particles we use the potential
    \begin{align}
        v(\vf{r})
        &=
        \begin{cases}
            \half m\omega^2r^2 & (\text{S}), \\
            \half m \bigl[
                \omega^2(x^2 + y^2) + \omega_z^2z^2
            \bigr] & (\text{E}),
        \end{cases}
    \end{align}
    where we can choose between a spherical (S) or an elliptical (E) harmonic
    oscillator trap. The full Hamiltonian of the system is given by
    \begin{align}
        H = \sum_{i = 1}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where the single particle one-body operator, $h(\vf{r}_i)$, is given by
    \begin{align}
        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
        + v(\vf{r}_i),
    \end{align}
    for equal masses. The two-body interaction operator, $w(\vf{r}_i,
    \vf{r}_j)$, is
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        = \begin{cases}
            \infty & |\vf{r}_i - \vf{r}_j| \leq a, \\
            0 & |\vf{r}_i - \vf{r}_j| > a,
        \end{cases}
        \label{eq:two-body_interaction}
    \end{align}
    where $a$ is the hard sphere of the particle.  The trial wavefunction we
    will be looking at is given by
    \begin{align}
        \Psi_T(\vf{r})
        &= \Phi_T(\vf{r})
        J(\vf{r}),
        %\\
        %&= \Biggl(
        %    \prod_{i = 1}^N g(\alpha, \beta, \vf{r}_i)
        %\Biggr)
        %\prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
        \label{eq:initial_trial_wavefunction}
    \end{align}
    where $\vf{r} = (\vf{r}_1, \dots, \vf{r}_N)$. The Slater permanent,
    $\Phi_T(\vf{r})$, is given by the $N$ first single-particle functions
    \begin{align}
        \Phi_T(\vf{r})
        &=
        \prod_{i = 1}^N \phi(\vf{r}_i)
        = \prod_{i = 1}^N\exp\bigl[
            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
        \bigr].
    \end{align}
    Here $\alpha$ is the variational parameter of the trial
    wavefunction.\footnote{We can also treat $\beta$ as a variational parameter,
    but we will in this project limit ourselves to a single variational
    parameter, i.e., $\alpha$.} The Jastrow factor, $J(\vf{r})$ contributes with
    correlations between the particles and is represented as
    \begin{align}
        J(\vf{r})
        &=
        \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
    \end{align}
    where the correlation wavefunctions are given by
    \begin{align}
        f(a, \vf{r}_j, \vf{r}_k)
        &=
        \begin{cases}
            0 & |\vf{r}_j - \vf{r}_k| \leq a, \\
            \Bigl(
                1 - \frac{a}{|\vf{r}_j - \vf{r}_k|}
            \Bigr) & |\vf{r}_j - \vf{r}_k| > a.
        \end{cases}
        \label{eq:correlation_wavefunction}
    \end{align}
    We will for brevity define the notation
    \begin{gather}
        \phi_i \equiv \phi(\vf{r}_i),
        \\
        r_{jk} \equiv \abs{\vf{r}_j - \vf{r}_k}.
    \end{gather}
    In practice \autoref{eq:two-body_interaction} is
    unnecessary as the situtation where $r_{jk} \leq a$ automatically yields a
    probability density of zero from \autoref{eq:correlation_wavefunction} thus
    rejecting the state completely.
